---
title: "Example plots - village profile"
author: "Nicholus"
date: '2024-01-23'

output: 
  beamer_presentation:
    keep_tex: true
header-includes:
  - \setbeamertemplate{footline}[frame number]
---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(stargazer)
library(magrittr)
library(haven)
library(foreign) # allows R to read .dta files from STATA

library(rio)
library(ggplot2)
library(survey)
library(reshape2)

```



# Village Characteristics

+ [Population](#pop) 
+ [Health facilities](#hfc) 
+ [Medicines and Equiptment Supply](#equiptment) 
+ [Main health concern](#healthconcern) 
+ [Covid-19 cases](#covid)
+ [Treated malnourished cases](#malnut) 
+ [Market information](#markets) 
+ [Food commodity prices](#foodprices) 
+ [Coping strategies](#vthccsi) 
+ [Telecom availability](#telecon) 


```{r, message = FALSE}

vthc <- rio::import("village.dta")

vthc <- vthc %>% 
  mutate(mother_svy = ifelse(mother_svy ==  1, "Mother Survey Villages", 
                             "No Mother Survey Villages"))

```

# Village Population (overall sample) {#pop}

```{r, message = FALSE, warning = F, fig.cap = "Village population distribution", fig.show='hold', out.width = '100%'}

mu <- vthc %>% summarise(grp.mean=mean(vill_pop_tot, na.rm = TRUE))

ggplot(vthc, aes(x=vill_pop_tot, fill = "white", color = "black")) +
  geom_histogram(position = "identity", alpha = 0.5)+
  geom_vline(data = mu, aes(xintercept = grp.mean, color = "black"),
             linetype="dashed")+
  theme(legend.position="none") +
    labs(x = "Village population", 
         y = "Number of Villages") +
  guides(fill = "none")

```


# Village Population (by sample type)


```{r, message = FALSE, warning = F, fig.cap = "Village Population Distribution per sample type", fig.show='hold', out.width = '100%'}

mu <- vthc %>% group_by(mother_svy) %>%
  summarise(grp.mean=mean(vill_pop_tot, na.rm = TRUE))


ggplot(vthc, aes(x=vill_pop_tot, fill = mother_svy, color=mother_svy)) +
  geom_histogram(position="identity", alpha=0.5)+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=mother_svy),
             linetype="dashed")+
  theme(legend.position="top") +
    labs(x = "Village population", 
         y = "Number of Villages", 
         color = "Survey village types") +
  guides(fill = "none")

```


# Village Population 

+ wider variation in village population with some outlier villages. 
+ villages where only VTHC surveys were collected, had higher population sizes.  


# Health Facilities (overall sample) {#hfc}

```{r, message = FALSE, warning = F, fig.cap = "Functioning health facilities (in Village Tract) distribution", fig.show='hold', out.width = '100%'}

hfc_level <- c("No Clinic", "Ethnic Health Organization Clinic", "Sub-Rural Health Center", 
               "Rural Health Center", "Public Hospital","Other Health Facility")


hfc <- melt(vthc, id.vars = 'mother_svy',
            measure.vars = c('hfc_type_0', 'hfc_type_1', 'hfc_type_2', 
                             'hfc_type_3', 'hfc_type_4', 'hfc_type_777')) %>%
  mutate(variable = recode(variable, 
         "hfc_type_0" = "No Clinic", 
         "hfc_type_1" = "Ethnic Health Organization Clinic", 
         "hfc_type_2" = "Sub-Rural Health Center", 
         "hfc_type_3" = "Rural Health Center", 
         "hfc_type_4" = "Public Hospital", 
         "hfc_type_777" = "Other Health Facility")) %>%
  mutate(variable = factor(variable, levels = hfc_level))

hfc %>% 
  group_by(variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = variable)) +
  geom_bar(position='dodge', stat='identity') + 
  geom_text(aes(label = prop), 
            size = 3,
            hjust = -.05) +
  labs(x = "Health Facility Types", 
       y = "Proportion of Villages") + 
  theme(legend.position = "none") +
  coord_flip()

```



# Health Facilities

+ almost half of the sample (40.3%) did not have access to the functioning clinic(s) 
+ among the accessible villages, EHO's clinics (19.5%) shared the largest proportion (by individual health facility type) 
+ but if we combined all the Government health facilities, its share was higher than EHO's clinics 


# Medicines and Equipment Supplies (overall sample) {#equiptment}

```{r, message = FALSE, warning = F, fig.cap = "Adequacy of medicines and equiptment supplires distribution", fig.show='hold', out.width = '100%'}

equip_level <- c("Adequate", "Not adequate", "Don't Know")

hfc_adequate <- melt(vthc, id.vars = 'mother_svy',
                     measure.vars = c("hfc_equipt_adequate_1", "hfc_equipt_adequate_2", 
                                      "hfc_equipt_adequate_888")) %>%
  mutate(variable = recode(variable, 
         "hfc_equipt_adequate_1" = "Adequate", 
         "hfc_equipt_adequate_2" = "Not adequate", 
         "hfc_equipt_adequate_888" = "Don't Know")) %>%
  mutate(variable = factor(variable, levels = equip_level))


hfc_adequate %>% 
  filter(!is.na(value)) %>%
  group_by(variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = variable)) +
  geom_bar(position='dodge', stat='identity') + 
  geom_text(aes(label = prop), 
            size = 3,
            hjust = -.05) +
  labs(x = "Medicines and Equipment Supplies", 
       y = "Proportion of Villages") + 
  theme(legend.position = "none")

```


# Main Health Concern (overall sample) {#healthconcern}

```{r, message = FALSE, warning = F, fig.cap = "Village main health concern as reported by VHTC", fig.show='hold', out.width = '100%'}

h_level <- c("Diarrhea", "Eye infections", "Vomiting", "Dehydration", "Snake bites", 
             "Fever", "Cough and fever (ARI)", "Skin disease", "Injuries/trauma", 
             "Mother or Child death following delivery", "Chronic diseases", 
             "Psychosocial illness", "Other health concerns")

h_issue <- melt(vthc, id.vars = 'mother_svy',
                measure.vars = c("health_issue_1", "health_issue_2", "health_issue_3", 
                                 "health_issue_4", "health_issue_5", "health_issue_6", 
                                 "health_issue_7", "health_issue_8", "health_issue_9", 
                                 "health_issue_10", "health_issue_11", "health_issue_12",
                                 "health_issue_777" )) %>%
  mutate(variable = recode(variable, 
         "health_issue_1" = "Diarrhea", 
         "health_issue_2" = "Eye infections", 
         "health_issue_3" = "Vomiting", 
         "health_issue_4" = "Dehydration", 
         "health_issue_5" = "Snake bites", 
         "health_issue_6" = "Fever", 
         "health_issue_7" = "Cough and fever (ARI)", 
         "health_issue_8" = "Skin disease", 
         "health_issue_9" = "Injuries/trauma", 
         "health_issue_10" = "Mother or Child death following delivery", 
         "health_issue_11" = "Chronic diseases", 
         "health_issue_12" = "Psychosocial illness", 
         "health_issue_777" = "Other health concerns")) %>%
  mutate(variable = factor(variable, levels = h_level))


h_issue %>% 
  group_by(variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = variable)) +
  geom_bar(position='dodge', stat='identity') + 
  geom_text(aes(label = prop), 
            size = 3,
            hjust = -.05) +
  labs(x = "Health Issues", 
       y = "Proportion of Villages") + 
  theme(legend.position = "none") +
  coord_flip()

```


# Main Health Concern (by sample type)

```{r, message = FALSE, warning = F, fig.cap = "Village main health concern by sample type (as reported by VHTC)", fig.show='hold', out.width = '100%'}

h_issue %>% 
  group_by(mother_svy, variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = mother_svy)) +
  geom_bar(position='dodge', stat='identity') + 
  labs(x = "Health Issues", 
       y = "Proportion of Villages", 
       fill = "Village Types") + 
  theme(legend.position = "bottom") +
  coord_flip()

```


# Main Health Concern 

+ common childhood illnesses like fever (79.2%) and cough (ARI) (55.8%) were the most reported diseases  
+ almost half of the village reported non-communicable disease(s) (41.6%)  
+ although the proportion was not higher than other common childhood illnesses, diarrhea was reported in a noticeable amount (around one out of three - 28.6%)  
+ mother survey (only) villages had a higher proportion of above-reported diseases, and among them, cough (ARI) was statistically significant 


# Covid-19 Suspected Cases {#covid}

```{r, message = FALSE, warning = F, fig.cap = "Covid-19 suspected cases distribution", fig.show='hold', out.width = '100%'}

vthc %>% 
  filter(!is.na(vill_covid_suspected)) %>%
  ggplot(aes(as.factor(vill_covid_suspected), ..count..)) +
  geom_bar(aes(fill = as.factor(vill_covid_suspected)), position = "dodge") +
  geom_text(stat = "count", 
            aes(label = ..count..), 
            size = 3,
            vjust = -.5) +
  labs(x = "Covid-19 Suspected Cases ", 
       y = "Number of Villages") +
  theme(legend.position = "none")

```


# Covid-19 Confirmed Cases 

```{r, message = FALSE, warning = F, fig.cap = "Covid-19 confirmed cases distribution", fig.show='hold', out.width = '100%'}

vthc %>% 
  filter(!is.na(vill_covid_confirmed)) %>%
  ggplot(aes(as.factor(vill_covid_confirmed), ..count..)) +
  geom_bar(aes(fill = as.factor(vill_covid_confirmed)), position = "dodge") +
  geom_text(stat = "count", 
            aes(label = ..count..), 
            size = 3,
            vjust = -.5) +
  labs(x = "Covid-19 Confirmed Cases", 
       y = "Number of Villages") +
  theme(legend.position = "none")
```


# Covid-19 Cases 

+ majority of villages did not have suspected (33 out of 77) or confirmed cases (55out of 77) 
+ wide variation of Covid-19 cases across different villages 



# Treated Malnutrition Cases (overall sample) {#malnut}

```{r, message = FALSE, warning = F, fig.cap = "Treated malnourished cases distribution", fig.show='hold', out.width = '100%'}

vthc %>% 
  filter(!is.na(hfc_malnutrition_case)) %>%
  ggplot(aes(as.factor(hfc_malnutrition_case), ..count..)) +
  geom_bar(aes(fill = as.factor(hfc_malnutrition_case)), position = "dodge") +
  geom_text(stat = "count", 
            aes(label = ..count..), 
            size = 3,
            vjust = -.5) +
  labs(x = "Treated Malnourished Cases", 
       y = "Number of Villages") +
  theme(legend.position = "none")

```


# Treated Malnutrition Cases (by sample type)

```{r, message = FALSE, warning = F, fig.cap = "Treated malnourished cases distribution by sample type", fig.show='hold', out.width = '100%'}

vthc %>% 
  filter(!is.na(hfc_malnutrition_case)) %>%
  ggplot(aes(as.factor(hfc_malnutrition_case), ..count..)) +
  geom_bar(aes(fill = as.factor(hfc_malnutrition_case)), position = "dodge") +
  geom_text(stat = "count", 
            aes(label = ..count..), 
            size = 3,
            vjust = -.5) +
  labs(x = "Treated Malnourished Cases", 
       y = "Number of Villages") +
  theme(legend.position = "none") +
  facet_wrap(vars(mother_svy), ncol = 2)

```


# Treated Malnutrition Cases 

+ majority of villages (55 out of 77) did not have treated malnourished cases. 
+ but this did not tell about the number of undernourished cases from each surveyed village (as it was not covered in the survey questionnaire). 

# Coping Strategies (overall sample) {#vthccsi}

```{r, message = FALSE, warning = F, fig.cap = "Common coping strategies distribution", fig.show='hold', out.width = '100%'}

cope_level <- c("Reduce food intake", "Eating low-quality food", "Increase borrowing", 
                "Sold HH assets", "Sold productive assets", "Migration", 
                "Reliance on external support", "Use of grain bank", 
                "Community helping each other", "Reliance on own farm's products", 
                "Other strategies", "No strategy available")

cope <- melt(vthc, id.vars = 'mother_svy',
             measure.vars = c("cope_practice_1", "cope_practice_2", "cope_practice_3", 
                              "cope_practice_4", "cope_practice_5", "cope_practice_6", 
                              "cope_practice_7", "cope_practice_8", "cope_practice_10", 
                              'cope_practice_11', "cope_practice_777", "cope_practice_9")) %>%
  mutate(variable = recode(variable, 
         "cope_practice_1" = "Reduce food intake", 
         "cope_practice_2" = "Eating low-quality food", 
         "cope_practice_3" = "Increase borrowing", 
         "cope_practice_4" = "Sold HH assets", 
         "cope_practice_5" = "Sold productive assets", 
         "cope_practice_6" = "Migration", 
         "cope_practice_7" = "Reliance on external support", 
         "cope_practice_8" = "Use of grain bank", 
         "cope_practice_10" = "Community helping each other", 
         "cope_practice_9" = "No strategy available", 
         "cope_practice_11" = "Reliance on own farm's products",
         "cope_practice_777" = "Other strategies")) %>%
  mutate(variable = factor(variable, levels = cope_level))

cope %>% 
  group_by(variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = variable)) +
  geom_bar(position='dodge', stat='identity') + 
  geom_text(aes(label = prop), 
            size = 3,
            hjust = -.05) +
  labs(x = "Type of Coping Strategies", 
       y = "Proportion of Villages") + 
  theme(legend.position = "none") +
  coord_flip()

```


# Coping Strategies 

+ Increasing borrowing (24.7%), selling productive assets (22.1%), and reduced food intakes (18.2%) were the most reported categories   
+ one-fifth of the sample reported other strategies (22.1%), and most of their answers were that they had not experienced the condition which required applying the coping mechanism 

# Distance to Market (overall sample) {#markets}


```{r, message = FALSE, warning = F, fig.cap = "Walking distance to market distribution", fig.show='hold', out.width = '100%'}

vthc %>%
  ggplot(aes("", market_walking, fill = "blue")) + 
  geom_boxplot(width=0.6) +
  stat_summary(
    aes(label=sprintf("%1.0f", ..y..)),
    geom = "text", 
    fun.y = function(y) boxplot.stats(y)$stats,
    position = position_nudge(x = 0.4), 
    size = 3) +
    labs(y = "Distance to Market", 
       x = "Duration (in minutes)") +
  theme(legend.position = "none") +
  coord_flip()

```


# Distance to Market (by sample type)

```{r, message = FALSE, warning = F, fig.cap = "Walking distance to market distribution by sample type", fig.show='hold', out.width = '100%'}

mu <- vthc %>%
  group_by(mother_svy) %>%
  summarise(grp.mean=mean(market_walking, na.rm = TRUE))


ggplot(vthc, aes(x=market_walking, fill = mother_svy, color=mother_svy)) +
  geom_histogram(position="identity", alpha=0.5)+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=mother_svy),
             linetype="dashed")+
  theme(legend.position="top") +
    labs(x = "Walking Distance to Markets (in minute)", 
         y = "Number of Villages", 
         color = "Survey village types") +
  guides(fill = "none")

```


# Distance to Market 

+ more villages required walking more than 2 hours (120 minutes)  
+ travel distances widely varied across villages, and outlier villages were identified (with over 12 hours of walking distance)  
+ mother survey villages required more travel time to access the market (median value -  2 hours), while the VTHC only surveyed village did 1:30 hours  


# Food commodity prices (overall sample) {#foodprices}

```{r, message = FALSE, warning = F, fig.cap = "Food commodity price distribution", fig.show='hold', out.width = '100%'}

price <- melt(vthc, id.vars = 'mother_svy',
            measure.vars = c("rice_price", "oil_price", "pulses_price", 
                             "meat_price", "potatoes_price" )) %>%
  mutate(variable = recode(variable, 
                           "rice_price" = "Rice", 
                           "oil_price" = "Oil",
                           "pulses_price" = "Pulses",
                           "meat_price" = "Meat", 
                           "potatoes_price" = "Potatoes"))

price %>%
  ggplot(aes(variable, value, fill = variable)) + 
  geom_boxplot(width=0.6) +
  stat_summary(
    aes(label=sprintf("%1.0f", ..y..)),
    geom = "text", 
    fun.y = median,
    position = position_nudge(x = 0.4), 
    size = 3) +
  labs(y = "Price", 
       x = "Type of Food") +
  theme(legend.position = "none") +
  coord_flip()
  
```



# Food commodity prices (by sample type)

```{r, message = FALSE, warning = F, fig.cap = "Food commodity price distribution by sample type", fig.show='hold', out.width = '100%'}

price %>%
  ggplot(aes(variable, value, fill = mother_svy)) + 
  geom_boxplot(width=0.6) +
  labs(y = "Price", 
       x = "Type of Food", 
       fill = "Survey village type") +
  theme(legend.position = "bottom") +
  coord_flip()
  
```

# Food commodity prices 

+ wide price variations were detected in meat and pulses food groups   
+ although the price had minor price variation compared to the above food groups, more outlier prices were seen, and more villages had higher prices than their median price value  
+ mother survey villages had a higher price for rice and potatoes but a lower oil price (statistically significant)  


# Telecom Availability (overall sample) {telecon}


```{r, message = FALSE, warning = F, fig.cap = "Telecom availability distribution", fig.show='hold', out.width = '100%'}

tele_level <- c('Radio', 'TV', 'Mobile phone', 'Internet', 'Other types')

telecon <- melt(vthc, id.vars = 'mother_svy',
             measure.vars = c("vill_telecoms_1", "vill_telecoms_2", "vill_telecoms_3",
                              "vill_telecoms_4", "vill_telecoms_777")) %>%
  mutate(variable = recode(variable, 
         "vill_telecoms_1" = 'Radio', 
         "vill_telecoms_2" = 'TV', 
         "vill_telecoms_3" = 'Mobile phone', 
         "vill_telecoms_4" = 'Internet', 
         "vill_telecoms_777" = 'Other types')) %>%
  mutate(variable = factor(variable, levels = tele_level))


telecon %>% 
  group_by(variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = variable)) +
  geom_bar(position='dodge', stat='identity') + 
  geom_text(aes(label = prop), 
            size = 2.5,
            hjust = -.05) +
  labs(x = "Telecom Type", 
       y = "Proportion of Villages") + 
  theme(legend.position = "none") +
  coord_flip()

```


# Telecom Availability (by sample type)

```{r, message = FALSE, warning = F, fig.cap = "Telecom availability distribution by sample type", fig.show='hold', out.width = '100%'}

telecon %>% 
  group_by(mother_svy, variable) %>%
  mutate(prop = round(sum(value)/n() * 100, 1)) %>%
  slice(1) %>%
  ungroup() %>%
  ggplot(aes(variable, prop, fill = mother_svy)) +
  geom_bar(position='dodge', stat='identity') + 
  labs(x = "Telecom Type", 
       y = "Proportion of Villages", 
       fill = "Village Types") + 
  theme(legend.position = "bottom") +
  coord_flip()

```

# Telecom Availability 

+ almost all surveyed villages (96.1%) had mobile phone access, but just over two-thirds of the villages got access to the internet 
+ majority of villages had the better coverage of radio and TV (over 70%) [means access to TV and radio, not all HHs from each village had radio or TV) 
+ only VTHC survey villages had the better condition in telecom accessibility  

