---
title: "r_int_ggplot"
author: "Nicholus Tint Zaw"
date: "2024-01-23"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
```


# Reference

reference: ggplot2: Elegant Graphics for Data Analysis (3e)
https://ggplot2-book.org/introduction 


# What is the grammar of graphics?

how the data’s variables are mapped to aesthetic attributes 

1. Layer: collection of geometric elements and statistical transformations  

    - Geometric elements, geoms for short, represent what you actually see in the plot: points, lines, polygons, etc.  
    - Statistical transformations, stats for short, summarise the data    
  
2. Scales: map values in the data space to values in the aesthetic space. This includes the use of colour, shape or size
3. Coord: coordinate system, describes how data coordinates are mapped to the plane of the graphic
4. Facet: how to break up and display subsets of data as small multiples
5. Theme: controls the finer points of display, like the font size and background colour


# Build a plot layer by layer


## Building a plot 

```{r}
p <- ggplot(mpg, aes(displ, hwy))
p

p + geom_point()

```


```{r}


mpg %>% 
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class))


```



```{r}

ggplot(mpg, aes(trans, cty)) + 
  geom_point() + 
  geom_point(stat = "summary", fun = "mean", colour = "red", size = 4) 



ggplot(mpg, aes(trans, cty)) + 
  geom_point() + 
  geom_point(stat = "summary", fun = "mean", colour = "red", size = 2) +
  geom_point(stat = "summary", fun = "median", colour = "blue", size = 2)


ggplot(mpg, aes(trans, cty)) + 
  geom_point() + 
  geom_point(stat = "summary", fun = "mean", colour = "red", size = 2) +
  geom_point(stat = "summary", fun = "median", colour = "blue", size = 2) + 
  coord_flip()




```


```{r, fig.height= 10, fig.width = 7}
ggplot(mpg, aes(trans, cty)) + 
  geom_point() + 
  geom_point(stat = "summary", fun = "mean", colour = "red", size = 2) +
  geom_point(stat = "summary", fun = "median", colour = "blue", size = 2) + 
  coord_flip() +
  facet_wrap(~class, ncol = 2)
```



